# Loading in RDS Phyloseq files (this part of the code by Abby Schaefer)

```{r}
source('http://bioconductor.org/biocLite.R')
biocLite('phyloseq')

#If bioconductor wont work in your environment you can use BiocManager as shown below to instal the phyloseq package
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install(c("phyloseq"))


setwd("./Apples/") #set this to the location you downloaded the apples.RDS file. Otherwise this won't work!
library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyverse)

seeds <- readRDS("./PhyloObj/seeds.RDS")
pulp <- readRDS("./PhyloObj/fruit_pulp.RDS")
calyx <- readRDS("./PhyloObj/calyx_end.RDS")
peel <- readRDS("./PhyloObj/peel.RDS")
stem <- readRDS("./PhyloObj/Stem.RDS")
stem_end <- readRDS("./PhyloObj/Stem_end.RDS")
apples <- merge_phyloseq(stem,stem_end,peel,pulp,seeds,calyx)

```


# Found way to more accurately recreate Figure 1 (With Abby's help!)
```{r}


applesdf <- psmelt(apples)
applesdf
ggplot() + geom_bar(aes(y= Abundance,x=tissue, fill= tissue), data=applesdf, stat= "identity")

```


# Counting Number of Reads (the initial plan to mimic Figure 1)

**Reading the OTU table from the phyloseq object**
```{r}
otu_table_copy <- apples@otu_table
```

**getting lists of sum of reads and sample names**
```{r}
#count number of rows to figure out range
rows <- nrow(otu_table_copy)

#initialize empty lists
reads <- c()
sample_name <- c()

#fill the lists by reading each row's sum and name
for (i in 1:nrow(otu_table_copy)) {
  total_reads <- sum(otu_table_copy[i])
  sample <- row.names(otu_table_copy[i])
  reads[i] <- total_reads
  sample_name[i] <- sample
}
```


**Create data frame from lists**
```{r}
reads_count.data <- data.frame(sample_name,reads)
```


**Separating out conventional management data  to make graph less crowded**

_Conventional_
```{r}
#grep only the conventional samples
conventional <- reads_count.data[grep(".C", reads_count.data$sample_name), ]

#make a number code to make the tickmarks readable
conventional <- mutate(conventional, num_code = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24))

#add sample type column so they can be filled according to type
conventional <- mutate(conventional, sample_type = c("Stem","Stem","Stem","Stem","Stem end","Stem end", "Stem end", "Stem end", "Peel", "Peel","Peel","Peel","Fruit pulp","Fruit pulp","Fruit pulp", "Fruit pulp", "Seed","Seed","Seed","Seed","Calyx end","Calyx end","Calyx end", "Calyx end"))
```

_Organic_
```{r}
#grep only the organic data
organic <- reads_count.data[grep(".H", reads_count.data$sample_name), ]

#create a number code for each sample to clean up the x axis
organic <- mutate(organic, num_code = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23,24))

#add column that has the type of tissue the sample is from.
organic <- mutate(organic, sample_type = c("Stem","Stem","Stem","Stem","Stem end","Stem end", "Stem end", "Stem end", "Peel", "Peel","Peel","Peel","Fruit pulp","Fruit pulp","Fruit pulp", "Fruit pulp", "Seed","Seed","Seed","Seed","Calyx end","Calyx end","Calyx end", "Calyx end"))
```

**Graph of Conventional samples**
```{r}
#make a bar graph of the conventional data where the x axis is the sample number and the y axis is the number of reads. Colored according to the sample type.
ggplot(data=conventional, aes(x=num_code, y=reads,fill=sample_type)) + ylab("Number of Reads")+ xlab("Sample Number")+ ggtitle("Conventional Apples")+
  geom_bar(stat="identity")
```

**Graph of Organic Samples**
```{r}
#make a bar graph of the organic data where the x axis is the sample number and the y axis is the number of reads. Colored according to the sample type.
ggplot(data=organic, aes(x=num_code, y=reads,fill=sample_type)) +ylab("Number of Reads")+ xlab("Sample Number")+ ggtitle("Organic Apples")+
  geom_bar(stat="identity")
```



