---
title: "Remaining Tissues"
author: "ClayBae"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library("knitr")
library("BiocStyle")
library("dada2")
library("phyloseq")
library("DECIPHER")
library("phangorn")
.cran_packages <- c("ggplot2", "gridExtra")
.bioc_packages <- c("dada2", "phyloseq", "DECIPHER", "phangorn")
.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
   install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
#if(any(!.inst)) {
 #  source("http://bioconductor.org/biocLite.R")
 #  biocLite(.bioc_packages[!.inst], ask = F)
#}
# Load packages into session, and print package version
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
```


```{r}
path <- "U:/Documents/ISU Classes/BCB 546X/ena_files"   #Change to the directory where the fastq files are after unzipping
list.files(path)
```

Next I read in the files that contain the reads from just the apple seeds samples. I renamed the files from the default names the ENA provided to have the suffix "_seeds" to make them easier to separate from the rst of the fastq files.
```{r}
seeds <- sort(list.files(path, pattern="_seed.fastq", full.names = TRUE))
sample.names <- sapply(strsplit(basename(seeds), "_"), `[`, 1)
```

Next I plotted the quality scores for each sample.
```{r}
plotQualityProfile(seeds[1:8])
```

I set up the object to place the filtered reads in.
```{r}
filtSeeds <- file.path(path, "filtered", paste0(sample.names, "filt.fastq"))
names(filtSeeds) <- sample.names
```

Next I filtered the reads using the default settings.
```{r}
out <- filterAndTrim(seeds, filtSeeds, truncQ = 2, truncLen = 0, maxLen = Inf, rm.phix=TRUE,
              compress=TRUE, multithread=FALSE) # On Windows set multithread=FALSE
head(out)
```

Then we learn the error rates of basecalling in the reads.
```{r}
errF <- learnErrors(filtSeeds, multithread=FALSE)
```

We plot the error rates to examine them.
```{r}
plotErrors(errF, nominalQ=TRUE)
```

Now we are ready to align the reads to reference sequences.
```{r}
dadaSeeds <- dada(filtSeeds, err=errF, multithread=TRUE)
```

```{r}
seqtab <- makeSequenceTable(dadaSeeds)
dim(seqtab)
```
```{r}
table(nchar(getSequences(seqtab)))
```

Remove chimeras.
```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=FALSE, verbose=TRUE)
```

And assign taxonomy. This could have been done using species level as well.
```{r}
taxa <- assignTaxonomy(seqtab.nochim, "U:/Documents/ISU Classes/BCB 546X/ena_files/silva_nr_v132_train_set.fa.gz", multithread=FALSE)
```

```{r}
samples.out <- rownames(seqtab.nochim)
library(dplyr)
```

Next I created the metadata.
```{r}
samples.out <- as.data.frame(samples.out)
samples.out <- mutate(samples.out, tissue = c('seed', 'seed', 'seed', 'seed', 'seed', 'seed', 'seed', 'seed'))
samples.out <- mutate(samples.out, mngmt = c('conventional', 'conventional', 'conventional', 'conventional', 'organic', 'organic', 'organic', 'organic'))
rownames(samples.out) <- samples.out$samples.out
samples.out$samples.out <- NULL
```

And finally generated the phyloseq object.
```{r}
apples <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows = FALSE), sample_data(samples.out), tax_table(taxa))
```

```{r}
dna <- Biostrings::DNAStringSet(taxa_names(apples))
names(dna) <- taxa_names(apples)
apples <- merge_phyloseq(apples, dna)
taxa_names(apples) <- paste0("ASV", seq(ntaxa(apples)))
```

```{r}
saveRDS(apples, "../apples.RDS")
```
